---
title: "Template for t-Test"
author: ""
date: ''
output: 
  html_document:
    toc: true
    toc_depth: 3
editor_options: 
  chunk_output_type: inline
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(dplyr)
library(haven)
library(moments)
library(ggpubr)
df <- read_sav("") #directory
```

# Testing Assumptions

## Summary statistics

### Template

```{r}
group_by(df, group) %>%
  summarise(
    count = n(),
    
    mean = unlist(lapply(list(variable1_name = variable1, 
                              variable2_name = variable2), 
                         mean)),
    
    sd = unlist(lapply(list(variable1, 
                            variable2),  
                       sd)),
    
    skwness = unlist(lapply(list(
    variable1,
    variable2),
    skewness)),
    
    kurtsis = unlist(lapply(list(
    variable1,
    variable2),
    kurtosis)),
    
    variables = c(names(mean))
  )
```

### Example

```{r}
group_by(df, Country) %>%
  summarise(
    count = n(),
    
    mean = unlist(lapply(list(Seeking_Help = Econ_HLP, 
                              Investing = Econ_INV, 
                              Additional_Work = Econ_WRK,
                              Traditional_Value = Econ_TVS,
                              Saving = Econ_SVE), 
                         mean)),
    
    sd = unlist(lapply(list(Econ_HLP, 
                            Econ_INV, 
                            Econ_WRK,
                            Econ_TVS,
                            Econ_SVE), 
                       sd)),
    
    skwness = unlist(lapply(list(
    Econ_HLP,
    Econ_INV,
    Econ_WRK,
    Econ_TVS,
    Econ_SVE),
    skewness)),
    
    kurtsis = unlist(lapply(list(
    Econ_HLP,
    Econ_INV,
    Econ_WRK,
    Econ_TVS,
    Econ_SVE),
    kurtosis)),
    
    variables = c(names(mean))
  )
```

skewness (and kurtosis) without dplyr:

```{r}
skwness = unlist(lapply(list(
    df$variable[which(df$group == "group1")], 
    df$variable[which(df$Country == "group2")]
    ), 
    skewness))
```

## Visualization

### Template

```{r}
ggboxplot(df, x = "group", y = c("variable1", "variable2"), 
          color = "group", palette = c("#00AFBB", "#E7B800"),
        ylab = c("variable1", "variable2"), xlab = "group")
```

### Example

```{r}

#box plot

ggboxplot(df, x = "Country", y = c("Econ_HLP", "Econ_INV", "Econ_WRK", "Econ_TVS", "Econ_SVE"), 
          color = "Country", palette = c("#00AFBB", "#E7B800"),
        ylab = c("Econ_HLP", "Econ_INV", "Econ_WRK", "Econ_TVS", "Econ_SVE"), xlab = "Country")

```

## Shapiro-Wilk Normality Test (N < 100)

sensitive. use with caution.

**for multiple variables:**

### Template

```{r}
# Shapiro-Wilk normality test for group 1

shapiro_g1 <- lapply(list(
            variable1_name = df$variable1[which(df$group == "group1")], 
            variable2_name = df$variable2[which(df$group == "group1")]), 
                         shapiro.test)

# Shapiro-Wilk normality test for group 2
shapiro_g1 <- lapply(list(
            variable1_name = df$variable1[which(df$group == "group2")], 
            variable2_name = df$variable2[which(df$group == "group2")]), 
                         shapiro.test)

#p-value should be greater than 0.05
```


### Example

```{r}
# Shapiro-Wilk normality test for group 1

shapiro_g1 <- lapply(list(
            Helping = df$Econ_HLP[which(df$Country == "Iran")], 
            Investing = df$Econ_INV[which(df$Country == "Iran")], 
            Additional_work = df$Econ_WRK[which(df$Country == "Iran")],
            Traditional_Value = df$Econ_TVS[which(df$Country == "Iran")],
            Saving = df$Econ_SVE[which(df$Country == "Iran")]), 
                         shapiro.test)

# Shapiro-Wilk normality test for group 2
shapiro_g2 <- lapply(list(
            Helping = df$Econ_HLP[which(df$Country == "Russia")], 
            Investing = df$Econ_INV[which(df$Country == "Russia")], 
            Additional_work = df$Econ_WRK[which(df$Country == "Russia")],
            Traditional_Value = df$Econ_TVS[which(df$Country == "Russia")],
            Saving = df$Econ_SVE[which(df$Country == "Russia")]), 
                         shapiro.test)

#p-value should be greater than 0.05
```

**for one variable:**

```{r}
# Shapiro-Wilk normality test for group 1
with(df, shapiro.test(variable1[group == "group1"]))

# Shapiro-Wilk normality test for group 2
with(df, shapiro.test(variable1[group == "group2"]))
```

## Two-sample Kolmogorov-Smirnov Normality Test (N > 100)

sensitive. use with caution.

**for multiple variables:**

### Template

```{r}
kstest <- lapply(list(variable1 ~ group, 
            variable2 ~ group), 
            ks.test, data = df)
```


### Example

```{r}
kstest <- lapply(list(Econ_HLP ~ Country, 
            Econ_INV ~ Country, 
            Econ_WRK ~ Country,
            Econ_TVS ~ Country,
            Econ_SVE ~ Country), 
                         ks.test, data = df)
```


**for one variable:**

```{r}
with(df, ks.test(variable1[group == "group1"], variable1[group == "group2"]))
```

## Homogeneity in Variances

**for multiple variables:**

### Template

```{r}
ftest <- lapply(list(variable1 ~ group, 
            variable2 ~ group), 
            var.test, data = df)
```

### Example

```{r}

ftest <- lapply(list(Econ_HLP ~ Country, 
            Econ_INV ~ Country, 
            Econ_WRK ~ Country,
            Econ_TVS ~ Country,
            Econ_SVE ~ Country), 
                         var.test, data = df)
ftest

#equal variances is assumed if p is greater than 0.05
```

**for one variable:**

```{r}
ftest <- var.test(variable1 ~ group, data = df)
ftest

#equal variances is assumed if p is greater than 0.05
```

# t-test

**for multiple variables:**

## Template

```{r}

result <- lapply(list(
            variable1 ~ group, 
            variable2 ~ group), 
                 t.test, 
                 data = df,
                 alternative = "two.sided", #other values: "greater", "less"
                 paired = FALSE, #choose whether it's independent sample or one sample t-test
                 var.equal = FALSE, #true if assumption of homoscedasticity is met
                 conf.level = .95
                 )
result
```

## Example

```{r}

result <- lapply(list(
            Econ_HLP ~ Country, 
            Econ_INV ~ Country, 
            Econ_WRK ~ Country,
            Econ_TVS ~ Country,
            Econ_SVE ~ Country), 
                 t.test, 
                 data = df,
                 alternative = "two.sided", #other values: "greater", "less"
                 paired = FALSE, #choose whether it's independent sample or one sample t-test
                 var.equal = FALSE, #true if assumption of homoscedasticity is met
                 conf.level = .95
                 )
result
```

**for one variable:**

```{r}
result <- t.test(variable1 ~ group, #numeric vectors, or fitted linear model objects (inheriting from class "lm").
       data = df,         
       alternative = "two.sided", #other values: "greater", "less"
       paired = FALSE, #choose whether it's independent sample or one sample t-test
       var.equal = FALSE, #true if assumption of homoscedasticity is met
       conf.level = .95,
       ) 
```

